---
title: "Prediction Assignment Writeup"
author: "ltl"
date: "Thursday, October 23, 2014"
output: html_document
---

### Introduction

In this document, I have used the data from Human Activity Recognition (HAR). the goal was to train a model to predict the manner(variable "classe") in which they did the exercise

The training data for this project are available here:

[https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv]

The test data are available here:

[https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv]

### reading and process the data

```{r, results='hide'}
setwd("D:\\study\\coursera\\8_Practical Machine Learning\\exercise\\project")
library(knitr);library(caret);library(randomForest)
opts_chunk$set( cache = TRUE )
```

- load data
- Remove columns with NA values
- Remove columns 1 to 7, which don't help in classifying data
- Remove near zero variance predictors, use function nearZeroVar, but "classe" is excluded 

```{r, results='hide'}
# load data
pml.training <- read.csv("pml-training.csv", header = TRUE, na.strings=c("NA",""))
pml.testing <- read.csv("pml-testing.csv", header = TRUE, na.strings=c("NA",""))
# Remove columns with NA values
colsel <- apply(pml.training,2,function(x) {sum(is.na(x))})
pml.training <- pml.training[, which(colsel == 0)]
pml.testing <- pml.testing[,  which(colsel == 0)]
# Remeve columns 1 to 7, because they don't help in classifying data
pml.training <- pml.training[, -c(1:7)]
pml.testing <- pml.testing[, -c(1:7)]
```

```{r}
# Remove near zero variance predictors, "classe" is excluded 
nearZeroVar(pml.training[,-53], saveMetrics = TRUE)$nzv
```

we can see all the columns' variance are not near zero, then we have 53 columns remain

### Data Partition

split the data to training and testing

```{r}
inTrain <- createDataPartition(y = pml.training$classe, p = 0.7, list = FALSE)
training <- pml.training[inTrain, ]
testing <- pml.training[-inTrain, ]
```

### random forest and cross validation

```{r}
# set cross validation in trainControl function
trainControl <- trainControl(method = "cv", number = 5) 
modFit <- train(classe ~ ., method = "rf", data = training, trControl = trainControl)
modFit$finalModel
```

### Predict model on the testing data and measure of out of sample error

we show the confusion matrix as follows, it contain the out of sample error:

```{r}
pred <- predict(modFit, newdata = testing)
confusionMatrix(pred, testing$classe)
```

### apply the model to the 20 test cases which has given in the question

```{r}
pred.test <- predict(modFit, newdata = pml.testing)
# show the result
pred.test
```

### output the prediction to 20 text file to answer the project 2

```{r}
# create output function
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
# output the file
pml_write_files(pred.test)
```